---
title: "Data Exploration"
author: "Emma Deering"
format: html
---

## Load in data
```{r}
library(readr)
clams <- read_csv("data/clams_clean.csv")
```

# Build growth plots 
```{r}
library(tidyverse)
library(ggplot2)

ggplot(data = clams, aes(x = growth_period, y = growth, group = name)) +
         geom_line() +
         facet_wrap(vars(location))
  
```

## Blacklake clams growth curves
```{r}
clams_blacklake <- clams |> filter(location == "BlackLake") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_blacklake, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_blacklake, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Briggs clams growth curves
```{r}
clams_briggs <- clams |> filter(location == "Briggs") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_briggs, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_briggs, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Colton Canoe Launch clams growth curves
```{r}
clams_colton <- clams |> filter(location == "ColtonCanoeLaunch") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_colton, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_colton, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Blacklake clams growth curves
```{r}
clams_coopersfalls <- clams |> filter(location == "CoopersFalls") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_coopersfalls, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_coopersfalls, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## DeKalb clams growth curves
```{r}
clams_dekalb <- clams |> filter(location == "DeKalb") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_dekalb, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_dekalb, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## EelWeir clams growth curves
```{r}
clams_eelweir <- clams |> filter(location == "EelWeir") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_eelweir, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_eelweir, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## GovPark clams growth curves
```{r}
clams_govpark <- clams |> filter(location == "GovPark") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_govpark, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_govpark, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Heuvelton Boat Launch clams growth curves
```{r}
clams_heuvelton <- clams |> filter(location == "HeuveltonBoatLaunch") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_heuvelton, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_heuvelton, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Horse Shoe Lake clams growth curves
```{r}
clams_horseshoe <- clams |> filter(location == "HorseShoeLake") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_horseshoe, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_horseshoe, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Lake Clear clams growth curves
```{r}
clams_clear <- clams |> filter(location == "LakeClear") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_clear, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_clear, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Lake Flower clams growth curves
```{r}
clams_flower <- clams |> filter(location == "LakeFlower") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_flower, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_flower, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


##Lehman Park clams growth curves
```{r}
clams_lehman <- clams |> filter(location == "LehamnPark") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_lehman, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_lehman, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


## Long Lake clams growth curves
```{r}
clams_long <- clams |> filter(location == "LongLake") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_long, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_long, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


## Massena clams growth curves
```{r}
clams_massena <- clams |> filter(location == "Massena") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_massena, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_massena, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


## Mirror Lake clams growth curves
```{r}
clams_mirror <- clams |> filter(location == "MirrorLake") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_mirror, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_mirror, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Mountain Lake clams growth curves
```{r}
clams_mountain <- clams |> filter(location == "MountainLake") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_mountain, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_mountain, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


## Norfolk clams growth curves
```{r}
clams_norfolk <- clams |> filter(location == "Norfolk") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_norfolk, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_norfolk, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


## Norfolk SM clams growth curves
```{r}
clams_norfolkSM <- clams |> filter(location == "NorfolkSM") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_norfolkSM, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_norfolkSM, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Norwood Beach clams growth curves
```{r}
clams_norwood <- clams |> filter(location == "NorwoodBeach") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_norwood, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_norwood, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Piercefield clams growth curves
```{r}
clams_piercefield <- clams |> filter(location == "Piercefield") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_piercefield, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_piercefield, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


## Pine Street clams growth curves
```{r}
clams_pine <- clams |> filter(location == "PineStreet") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_pine, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_pine, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Raquette Pond clams growth curves
```{r}
clams_raquette <- clams |> filter(location == "RaquettePond") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_raquette, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_raquette, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## South Colton Bridge clams growth curves
```{r}
clams_southcolton <- clams |> filter(location == "SouthColtonBridge") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_southcolton, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_southcolton, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Tupper Horse Shoe Lake clams growth curves
```{r}
clams_tupperhorseshoe <- clams |> filter(location == "TupperHorseShoeLake") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_tupperhorseshoe, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_tupperhorseshoe, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```


## Tupper Lake North clams growth curves
```{r}
clams_tuppernorth <- clams |> filter(location == "TupperLakeNorth") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_tuppernorth, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_tuppernorth, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```

## Union Falls Pond clams growth curves
```{r}
clams_union <- clams |> filter(location == "UnionFallsPond") |> group_by(name)

# Multiple Graphs
ggplot(data=clams_union, aes(x = growth_period, y = growth)) + 
  geom_line() +
  facet_wrap(vars(name))

# Single Graphs
ggplot(clams_union, aes(x = growth_period, y = growth, 
                            color = name, group = name)) + 
  geom_line()

```
