---
title: "03 - Individual Fits"
author: "Emma Deering"
format: html
---

Read in Data:
```{r}
library(readr)
clams <- read_csv("data/clams_clean.csv")
View(clams)
```

Remove negative values:
```{r}
library(dplyr)
library(ggplot2)
clams <- clams |> filter(growth > 0)

# removes negative values from Coopers Falls, Lake Clear, Mirror Lake
```

List of all the Water Bodies:
```{r}
clams |> group_by(location) |> summarise(min(growth_period))
```

Black Lake: 
```{r}
clams_blacklake <- clams |>
  filter(location == "BlackLake") |>
  group_by(name)

blacklake_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_blacklake,
  start = list(max_l = max(clams_blacklake$Length), k = 0.1, b = 0))

predict_blacklake <- clams_blacklake |>
  ungroup() |>
  mutate(predicted_length = predict(blacklake_vb)) 

blacklake <- ggplot(predict_blacklake,
                          aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "lightgreen", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of BLack Lake Clams")

blacklake
```

Briggs:
```{r}
clams_briggs <- clams |>
  filter(location == "Briggs") |>
  group_by(name)

briggs_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_briggs,
  start = list(max_l = max(clams_briggs$Length), k = 0.1, b = 0))

predict_briggs <- clams_briggs |>
  ungroup() |>
  mutate(predicted_length = predict(briggs_vb)) 

briggs <- ggplot(predict_briggs,
                          aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "green", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Briggs Clams")

briggs
```

Colton Canoe Launch:
```{r}
clams_colton <- clams |>
  filter(location == "ColtonCanoeLaunch") |>
  group_by(name)

colton_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_colton,
  start = list(max_l = max(clams_colton$Length), k = 0.1, b = 0))

predict_colton <- clams_colton |>
  ungroup() |>
  mutate(predicted_length = predict(colton_vb)) 

colton <- ggplot(predict_colton,
                          aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "magenta", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Colton Canoe Launch Clams")

colton
```

Coppers Falls:
```{r}
clams_copper <- clams |>
  filter(location == "CoppersFalls") |>
  group_by(name)

copper_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_copper,
  start = list(max_l = max(clams_copper$Length), k = 0.1, b = 0))

predict_copper <- clams_copper |>
  ungroup() |>
  mutate(predicted_length = predict(copper_vb)) 

copper <- ggplot(predict_copper,
                          aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "gold", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Coppers Falls Clams")

copper
```

De Kalb:
```{r}
clams_dekalb <- clams |>
  filter(location == "DeKalb") |>
  group_by(name)

dekalb_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_dekalb,
  start = list(max_l = max(clams_dekalb$Length), k = 0.1, b = 0))

predict_dekalb <- clams_dekalb |>
  ungroup() |>
  mutate(predicted_length = predict(dekalb_vb)) 

dekalb <- ggplot(predict_dekalb,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkgreen", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of De Kalb Clams")

dekalb

```

Eel Weir:
```{r}
clams_eelweir <- clams |>
  filter(location == "EelWeir") |>
  group_by(name)

eelweir_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_eelweir,
  start = list(max_l = max(clams_eelweir$Length), k = 0.1, b = 0))

predict_eelweir <- clams_eelweir |>
  ungroup() |>
  mutate(predicted_length = predict(eelweir_vb)) 

eelweir <- ggplot(predict_eelweir,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "blue", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Eel Weir Clams")

eelweir

```

Gov Park:
```{r}
clams_govpark <- clams |>
  filter(location == "GovPark") |>
  group_by(name)

govpark_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_govpark,
  start = list(max_l = max(clams_govpark$Length), k = 0.1, b = 0))

predict_govpark <- clams_govpark |>
  ungroup() |>
  mutate(predicted_length = predict(govpark_vb)) 

govpark <- ggplot(predict_govpark,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "purple", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Gov Park Clams")

govpark

```

Heuvelton Boat Launch:
```{r}
clams_heuvelton <- clams |>
  filter(location == "HeuveltonBoatLaunch") |>
  group_by(name)

heuvelton_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_heuvelton,
  start = list(max_l = max(clams_heuvelton$Length), k = 0.1, b = 0))

predict_heuvelton <- clams_heuvelton |>
  ungroup() |>
  mutate(predicted_length = predict(heuvelton_vb)) 

heuvelton <- ggplot(predict_heuvelton,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "orange", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Heuvelton Boat Launch Clams")

heuvelton

```

Horse Shoe Lake:
```{r}
clams_horseshoe <- clams |>
  filter(location == "HorseShoeLake") |>
  group_by(name)

horseshoe_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_horseshoe,
  start = list(max_l = max(clams_horseshoe$Length), k = 0.1, b = 0))

predict_horseshoe <- clams_horseshoe |>
  ungroup() |>
  mutate(predicted_length = predict(horseshoe_vb)) 

horseshoe <- ggplot(predict_horseshoe,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "brown", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Horse Shoe Lake Clams")

horseshoe

```

Lake Clear:
```{r}
clams_lakeclear <- clams |>
  filter(location == "LakeClear") |>
  group_by(name)

lakeclear_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_lakeclear,
  start = list(max_l = max(clams_lakeclear$Length), k = 0.1, b = 0))

predict_lakeclear <- clams_lakeclear |>
  ungroup() |>
  mutate(predicted_length = predict(lakeclear_vb)) 

lakeclear <- ggplot(predict_lakeclear,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "aquamarine", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Lake Clear Clams")

lakeclear

```

Lake Flower:
```{r}
clams_lakeflower <- clams |>
  filter(location == "LakeFlower") |>
  group_by(name)

lakeflower_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_lakeflower,
  start = list(max_l = max(clams_lakeflower$Length), k = 0.1, b = 0))

predict_lakeflower <- clams_lakeflower |>
  ungroup() |>
  mutate(predicted_length = predict(lakeflower_vb)) 

lakeflower <- ggplot(predict_lakeflower,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkseagreen", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Lake Flower Clams")

lakeflower

```

Lehman Park:
```{r}
clams_lehman <- clams |>
  filter(location == "LehmanPark") |>
  group_by(name)

lehman_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_lehman,
  start = list(max_l = max(clams_lehman$Length), k = 0.1, b = 0))

predict_lehman <- clams_lehman |>
  ungroup() |>
  mutate(predicted_length = predict(lehman_vb)) 

lehman <- ggplot(predict_lehman,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "steelblue", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Lehman Park Clams")

lehman

```

Long Lake:
```{r}
clams_longlake <- clams |>
  filter(location == "LongLake") |>
  group_by(name)

longlake_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_longlake,
  start = list(max_l = max(clams_longlake$Length), k = 0.1, b = 0))

predict_longlake <- clams_longlake |>
  ungroup() |>
  mutate(predicted_length = predict(longlake_vb)) 

longlake <- ggplot(predict_longlake,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkorchid", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Long Lake Clams")

longlake

```

Massena:
```{r}
clams_massena <- clams |>
  filter(location == "Massena") |>
  group_by(name)

massena_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_massena,
  start = list(max_l = max(clams_massena$Length), k = 0.1, b = 0))

predict_massena <- clams_massena |>
  ungroup() |>
  mutate(predicted_length = predict(massena_vb)) 

massena <- ggplot(predict_massena,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkorange", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Massena Clams")

massena

```

Mirror Lake:
```{r}
clams_mirror <- clams |>
  filter(location == "MirrorLake") |>
  group_by(name)

mirror_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_mirror,
  start = list(max_l = max(clams_mirror$Length), k = 0.1, b = 0))

predict_mirror <- clams_mirror |>
  ungroup() |>
  mutate(predicted_length = predict(mirror_vb)) 

mirror <- ggplot(predict_mirror,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "hotpink", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Mirror Lake Clams")

mirror

```

Mountain Lake:
```{r}
clams_mountain <- clams |>
  filter(location == "MountainLake") |>
  group_by(name)

mountain_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_mountain,
  start = list(max_l = max(clams_mountain$Length), k = 0.1, b = 0))

predict_mountain <- clams_mountain |>
  ungroup() |>
  mutate(predicted_length = predict(mountain_vb)) 

mountain <- ggplot(predict_mountain,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "navy", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Mountain Lake Clams")

mountain

```

Norfolk:
```{r}
clams_norfolk <- clams |>
  filter(location == "Norfolk") |>
  group_by(name)

norfolk_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_norfolk,
  start = list(max_l = max(clams_norfolk$Length), k = 0.1, b = 0))

predict_norfolk <- clams_norfolk |>
  ungroup() |>
  mutate(predicted_length = predict(norfolk_vb)) 

norfolk <- ggplot(predict_norfolk,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkcyan", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Norfolk Clams")

norfolk

```

Norfolk SM:
```{r}
clams_norfolksm <- clams |>
  filter(location == "NorfolkSM") |>
  group_by(name)

norfolksm_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_norfolksm,
  start = list(max_l = max(clams_norfolksm$Length), k = 0.1, b = 0))

predict_norfolksm <- clams_norfolksm |>
  ungroup() |>
  mutate(predicted_length = predict(norfolksm_vb)) 

norfolksm <- ggplot(predict_norfolksm,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "goldenrod", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Norfolk SM Clams")

norfolksm

```

Norwood Beach:
```{r}
clams_norwood <- clams |>
  filter(location == "NorwoodBeach") |>
  group_by(name)

norwood_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_norwood,
  start = list(max_l = max(clams_norwood$Length), k = 0.1, b = 0))

predict_norwood <- clams_norwood |>
  ungroup() |>
  mutate(predicted_length = predict(norwood_vb)) 

norwood <- ggplot(predict_norwood,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkmagenta", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Norwood Beach Clams")

norwood

```

Piercefield:
```{r}
clams_piercefield <- clams |>
  filter(location == "Piercefield") |>
  group_by(name)

piercefield_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_piercefield,
  start = list(max_l = max(clams_piercefield$Length), k = 0.1, b = 0))

predict_piercefield <- clams_piercefield |>
  ungroup() |>
  mutate(predicted_length = predict(piercefield_vb)) 

piercefield <- ggplot(predict_piercefield,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkslateblue", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Piercefield Clams")

piercefield

```

Pine Street:
```{r}
clams_pinestreet <- clams |>
  filter(location == "PineStreet") |>
  group_by(name)

pinestreet_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_pinestreet,
  start = list(max_l = max(clams_pinestreet$Length), k = 0.1, b = 0))

predict_pinestreet <- clams_pinestreet |>
  ungroup() |>
  mutate(predicted_length = predict(pinestreet_vb)) 

pinestreet <- ggplot(predict_pinestreet,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "forestgreen", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Pine Street Clams")

pinestreet

```

Raquette Pond:
```{r}
clams_raquette <- clams |>
  filter(location == "RaquettePond") |>
  group_by(name)

raquette_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_raquette,
  start = list(max_l = max(clams_raquette$Length), k = 0.1, b = 0))

predict_raquette <- clams_raquette |>
  ungroup() |>
  mutate(predicted_length = predict(raquette_vb)) 

raquette <- ggplot(predict_raquette,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "darkgoldenrod", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Raquette Pond Clams")

raquette

```

Tupper Horse Shoe Lake:
```{r}
clams_tupperhorseshoe <- clams |>
  filter(location == "TupperHorseShoeLake") |>
  group_by(name)

tupperhorseshoe_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_tupperhorseshoe,
  start = list(max_l = max(clams_tupperhorseshoe$Length), k = 0.1, b = 0))

predict_tupperhorseshoe <- clams_tupperhorseshoe |>
  ungroup() |>
  mutate(predicted_length = predict(tupperhorseshoe_vb)) 

tupperhorseshoe <- ggplot(predict_tupperhorseshoe,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "maroon3", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Tupper Horse Shoe Lake Clams")

tupperhorseshoe

```


Tupper Lake North:
```{r}
clams_tupperlakenorth <- clams |>
  filter(location == "TupperLakeNorth") |>
  group_by(name)

tupperlakenorth_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_tupperlakenorth,
  start = list(max_l = max(clams_tupperlakenorth$Length), k = 0.1, b = 0))

predict_tupperlakenorth <- clams_tupperlakenorth |>
  ungroup() |>
  mutate(predicted_length = predict(tupperlakenorth_vb)) 

tupperlakenorth <- ggplot(predict_tupperlakenorth,
                          aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "slateblue", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Tupper Lake North Clams")

tupperlakenorth

```

South Colton Bridge: 
```{r}
clams_southcoltonbridge <- clams |>
  filter(location == "SouthColtonBridge") |>
  group_by(name)

southcoltonbridge_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_southcoltonbridge,
  start = list(max_l = max(clams_southcoltonbridge$Length), k = 0.1, b = 0))

predict_southcoltonbridge <- clams_southcoltonbridge |>
  ungroup() |>
  mutate(predicted_length = predict(southcoltonbridge_vb)) 

southcoltonbridge <- ggplot(predict_southcoltonbridge,
                          aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "yellow", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of South Colton Bridge Clams")

southcoltonbridge
```

Union Falls Pond: 
```{r}
clams_unionfallspond <- clams |>
  filter(location == "UnionFallsPond") |>
  group_by(name)

unionfallspond_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_unionfallspond,
  start = list(max_l = max(clams_unionfallspond$Length), k = 0.1, b = 0))

predict_unionfallspond <- clams_unionfallspond |>
  ungroup() |>
  mutate(predicted_length = predict(unionfallspond_vb)) 

unionfallspond <- ggplot(predict_unionfallspond,
                          aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "mediumvioletred", linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length",
       title = "Growth Curve of Union Falls Pond Clams")

unionfallspond
```

________________________________________________________________________________________________

 Read in lakes and rivers data
```{r}
# Lakes and Rivers Data
library(readxl)
covariate_data <- read_excel("data/covariate_data.xlsx")
covariate_data
```

Divide by Lake/River:
```{r}
clams_lakes <- clams_total |> filter(body == "L")
clams_rivers <- clams_total |> filter(body == "R")
```


```{r}
# Bind growth predictions into one data set
predictions <- bind_rows(predict_unionfallspond, predict_southcoltonbridge, predict_tupperlakenorth, predict_tupperhorseshoe, predict_raquette, predict_pinestreet, predict_piercefield, predict_norwood, predict_norfolk, predict_norfolksm, predict_mountain, predict_mirror, predict_massena, predict_longlake, predict_lehman, predict_lakeclear, predict_lakeflower, predict_horseshoe, predict_heuvelton, predict_govpark, predict_eelweir, predict_dekalb, predict_copper, predict_colton, predict_briggs, predict_blacklake, predict_birggs)

predictions

# add with lake/river data
clams_predictions_waterbody <- predictions |> left_join(covariate_data, by = "location")

clams_predictions_waterbody
```

```{r}
# plot to compare between lakes and rivers

ggplot(data = clams_predictions_waterbody,  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length), linewidth = 2) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length", title = "Growth Curves by Lakes and Rivers") +
  facet_wrap(~body) + 
  geom_smooth()

```

It appears rivers have more standard growth compared to lakes. This is likely due to currents. clams are mostly filter feeders and are reliant on currents to bring them food. 