---
title: "Modeling Growth Rates of Fossilized Bivalves Using Bayesian Statistics"
author:
  - name: Emma Deering <br>
affiliation:
  - address: Department of Math, Computer Science, and Statistics <br> St. Lawrence University <br> Advised by Dr. Matt Higham
column_numbers: 3
output: 
  posterdown::posterdown_html:
   self_contained: true
poster_height: "36in"
poster_width: "48in"
titlebox_borderwidth: "0cm"
sectitle_textsize: "62pt"
primary_colour: "#49585C"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	singlespacing = TRUE
)
library(tidyverse)
library(pander)
library(dplyr)
library(ggplot2)
library(readr)
library(pander)
library(readxl)
```

# Abstract

Exploring how growth rates of fossilized clams change across environmental conditions across Northern New York, specifically in regards to acidic vs non acidic bodies of water.

**Key Objectives:**

-   Use a Von Bertanfly growth model to model clam growth rates.

-   Estimate model parameters using a MCMC algorithm.

-   Incorporate a Bayesian framework to account for uncertainty.

# Data

-   Using shell growth rings, the data records how much each clam grew during each annual growth period.

-   The data set contains the following useful variables:

    -   Location: The lake or river that the clam lived in.

    -   Growth Period: Indicates annual stage of growth.

    -   Length: Overall length of shell.

    -   Growth: Total growth during that growth period.

    -   Body: Indicates if water body was a lake or river.

-   Data collected Dr. Judith Nagel-Myers of the St. Lawrence University (Geology).

-   151 valves from 17 different localities.

# Growth Model and Parameters

Von Bertanfly growth model models clam growth as length versus age, capturing decreasing growth rates over time typical of organism development.

$$
L(t) = L \cdot (1 - e^{-k \cdot (t-t_0)})
$$ where

-   $t$ is time.

-   $L(t)$ is mean length at time t.

-   $max\_l$ is the asymptotic length.

-   $k$ is how quickly the clams reaches $max\_l$.

-   $b$ is time where length is 0.

```{r}
# sample von bertanfly growth model - simple

clams <- read_csv("data/clams_clean.csv")

# remove negative values from Coopers Falls, Lake Clear, Mirror Lake
clams <- clams |> filter(growth > 0)

clams_norfolksm <- clams |>
  filter(location == "NorfolkSM") |>
  group_by(name)

von_bert_function <- function(t, max_l, k, b) {
  max_l * (1 - exp(-k * (t - b)))}

norfolksm_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_norfolksm,
  start = list(max_l = max(clams_norfolksm$Length), k = 0.1, b = 0))

predict_norfolksm <- clams_norfolksm |>
  ungroup() |>
  mutate(predicted_length = predict(norfolksm_vb))

norfolksm <- ggplot(predict_norfolksm,
  aes(x = growth_period, y = Length)) +
  geom_line(aes(group = name)) +
  geom_line(aes(y = predicted_length),
            color = "darkblue", linewidth = 2) +
  labs(x = "Growth Period", y = "Length",
       title = "Von Bertanfly Growth Curve Example",
       caption = stringr::str_wrap("Blue line represents the general Von Bertanfly growth model for a collection of clams in a specific location. Black lines represent the growth of each clam in the locality.")) + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16), 
        plot.caption = element_text(
        size = 14,     
        hjust = 0.5))   
```

```{r}
# plot to compare between lakes and rivers

# data predictions
clams_blacklake <- clams |>
  filter(location == "BlackLake") |>
  group_by(name)

blacklake_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_blacklake,
  start = list(max_l = max(clams_blacklake$Length), k = 0.1, b = 0))

predict_blacklake <- clams_blacklake |>
  ungroup() |>
  mutate(predicted_length = predict(blacklake_vb)) 

clams_briggs <- clams |>
  filter(location == "Briggs") |>
  group_by(name)

briggs_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_briggs,
  start = list(max_l = max(clams_briggs$Length), k = 0.1, b = 0))

predict_briggs <- clams_briggs |>
  ungroup() |>
  mutate(predicted_length = predict(briggs_vb)) 

clams_colton <- clams |>
  filter(location == "ColtonCanoeLaunch") |>
  group_by(name)

colton_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_colton,
  start = list(max_l = max(clams_colton$Length), k = 0.1, b = 0))

predict_colton <- clams_colton |>
  ungroup() |>
  mutate(predicted_length = predict(colton_vb))

clams_copper <- clams |>
  filter(location == "CoppersFalls") |>
  group_by(name)

copper_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_copper,
  start = list(max_l = max(clams_copper$Length), k = 0.1, b = 0))

predict_copper <- clams_copper |>
  ungroup() |>
  mutate(predicted_length = predict(copper_vb))

clams_dekalb <- clams |>
  filter(location == "DeKalb") |>
  group_by(name)

dekalb_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_dekalb,
  start = list(max_l = max(clams_dekalb$Length), k = 0.1, b = 0))

predict_dekalb <- clams_dekalb |>
  ungroup() |>
  mutate(predicted_length = predict(dekalb_vb))

clams_eelweir <- clams |>
  filter(location == "EelWeir") |>
  group_by(name)

eelweir_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_eelweir,
  start = list(max_l = max(clams_eelweir$Length), k = 0.1, b = 0))

predict_eelweir <- clams_eelweir |>
  ungroup() |>
  mutate(predicted_length = predict(eelweir_vb)) 

clams_govpark <- clams |>
  filter(location == "GovPark") |>
  group_by(name)

govpark_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_govpark,
  start = list(max_l = max(clams_govpark$Length), k = 0.1, b = 0))

predict_govpark <- clams_govpark |>
  ungroup() |>
  mutate(predicted_length = predict(govpark_vb)) 

clams_heuvelton <- clams |>
  filter(location == "HeuveltonBoatLaunch") |>
  group_by(name)

heuvelton_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_heuvelton,
  start = list(max_l = max(clams_heuvelton$Length), k = 0.1, b = 0))

predict_heuvelton <- clams_heuvelton |>
  ungroup() |>
  mutate(predicted_length = predict(heuvelton_vb))

clams_horseshoe <- clams |>
  filter(location == "HorseShoeLake") |>
  group_by(name)

horseshoe_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_horseshoe,
  start = list(max_l = max(clams_horseshoe$Length), k = 0.1, b = 0))

predict_horseshoe <- clams_horseshoe |>
  ungroup() |>
  mutate(predicted_length = predict(horseshoe_vb)) 

clams_lakeclear <- clams |>
  filter(location == "LakeClear") |>
  group_by(name)

lakeclear_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_lakeclear,
  start = list(max_l = max(clams_lakeclear$Length), k = 0.1, b = 0))

predict_lakeclear <- clams_lakeclear |>
  ungroup() |>
  mutate(predicted_length = predict(lakeclear_vb)) 

clams_lakeflower <- clams |>
  filter(location == "LakeFlower") |>
  group_by(name)

lakeflower_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_lakeflower,
  start = list(max_l = max(clams_lakeflower$Length), k = 0.1, b = 0))

predict_lakeflower <- clams_lakeflower |>
  ungroup() |>
  mutate(predicted_length = predict(lakeflower_vb)) 

clams_lehman <- clams |>
  filter(location == "LehmanPark") |>
  group_by(name)

lehman_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_lehman,
  start = list(max_l = max(clams_lehman$Length), k = 0.1, b = 0))

predict_lehman <- clams_lehman |>
  ungroup() |>
  mutate(predicted_length = predict(lehman_vb)) 

clams_longlake <- clams |>
  filter(location == "LongLake") |>
  group_by(name)

longlake_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_longlake,
  start = list(max_l = max(clams_longlake$Length), k = 0.1, b = 0))

predict_longlake <- clams_longlake |>
  ungroup() |>
  mutate(predicted_length = predict(longlake_vb))

clams_massena <- clams |>
  filter(location == "Massena") |>
  group_by(name)

massena_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_massena,
  start = list(max_l = max(clams_massena$Length), k = 0.1, b = 0))

predict_massena <- clams_massena |>
  ungroup() |>
  mutate(predicted_length = predict(massena_vb)) 

clams_mirror <- clams |>
  filter(location == "MirrorLake") |>
  group_by(name)

mirror_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_mirror,
  start = list(max_l = max(clams_mirror$Length), k = 0.1, b = 0))

predict_mirror <- clams_mirror |>
  ungroup() |>
  mutate(predicted_length = predict(mirror_vb)) 

clams_mountain <- clams |>
  filter(location == "MountainLake") |>
  group_by(name)

mountain_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_mountain,
  start = list(max_l = max(clams_mountain$Length), k = 0.1, b = 0))

predict_mountain <- clams_mountain |>
  ungroup() |>
  mutate(predicted_length = predict(mountain_vb)) 

clams_norfolk <- clams |>
  filter(location == "Norfolk") |>
  group_by(name)

norfolk_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_norfolk,
  start = list(max_l = max(clams_norfolk$Length), k = 0.1, b = 0))

predict_norfolk <- clams_norfolk |>
  ungroup() |>
  mutate(predicted_length = predict(norfolk_vb)) 

clams_norfolksm <- clams |>
  filter(location == "NorfolkSM") |>
  group_by(name)

norfolksm_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_norfolksm,
  start = list(max_l = max(clams_norfolksm$Length), k = 0.1, b = 0))

predict_norfolksm <- clams_norfolksm |>
  ungroup() |>
  mutate(predicted_length = predict(norfolksm_vb)) 

clams_norwood <- clams |>
  filter(location == "NorwoodBeach") |>
  group_by(name)

norwood_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_norwood,
  start = list(max_l = max(clams_norwood$Length), k = 0.1, b = 0))

predict_norwood <- clams_norwood |>
  ungroup() |>
  mutate(predicted_length = predict(norwood_vb))

clams_piercefield <- clams |>
  filter(location == "Piercefield") |>
  group_by(name)

piercefield_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_piercefield,
  start = list(max_l = max(clams_piercefield$Length), k = 0.1, b = 0))

predict_piercefield <- clams_piercefield |>
  ungroup() |>
  mutate(predicted_length = predict(piercefield_vb)) 

clams_pinestreet <- clams |>
  filter(location == "PineStreet") |>
  group_by(name)

pinestreet_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_pinestreet,
  start = list(max_l = max(clams_pinestreet$Length), k = 0.1, b = 0))

predict_pinestreet <- clams_pinestreet |>
  ungroup() |>
  mutate(predicted_length = predict(pinestreet_vb)) 

clams_raquette <- clams |>
  filter(location == "RaquettePond") |>
  group_by(name)

raquette_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_raquette,
  start = list(max_l = max(clams_raquette$Length), k = 0.1, b = 0))

predict_raquette <- clams_raquette |>
  ungroup() |>
  mutate(predicted_length = predict(raquette_vb)) 

clams_unionfallspond <- clams |>
  filter(location == "UnionFallsPond") |>
  group_by(name)

unionfallspond_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_unionfallspond,
  start = list(max_l = max(clams_unionfallspond$Length), k = 0.1, b = 0))

predict_unionfallspond <- clams_unionfallspond |>
  ungroup() |>
  mutate(predicted_length = predict(unionfallspond_vb)) 

clams_southcoltonbridge <- clams |>
  filter(location == "SouthColtonBridge") |>
  group_by(name)

southcoltonbridge_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_southcoltonbridge,
  start = list(max_l = max(clams_southcoltonbridge$Length), k = 0.1, b = 0))

predict_southcoltonbridge <- clams_southcoltonbridge |>
  ungroup() |>
  mutate(predicted_length = predict(southcoltonbridge_vb)) 

clams_tupperlakenorth <- clams |>
  filter(location == "TupperLakeNorth") |>
  group_by(name)

tupperlakenorth_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_tupperlakenorth,
  start = list(max_l = max(clams_tupperlakenorth$Length), k = 0.1, b = 0))

predict_tupperlakenorth <- clams_tupperlakenorth |>
  ungroup() |>
  mutate(predicted_length = predict(tupperlakenorth_vb)) 

clams_tupperhorseshoe <- clams |>
  filter(location == "TupperHorseShoeLake") |>
  group_by(name)

tupperhorseshoe_vb <- nls(Length ~ von_bert_function(growth_period, max_l, k, b),
  data = clams_tupperhorseshoe,
  start = list(max_l = max(clams_tupperhorseshoe$Length), k = 0.1, b = 0))

predict_tupperhorseshoe <- clams_tupperhorseshoe |>
  ungroup() |>
  mutate(predicted_length = predict(tupperhorseshoe_vb)) 

# covariate data
covariate_data <- read_excel("data/covariate_data.xlsx")

# Bind growth predictions into one data set
predictions <- bind_rows(predict_unionfallspond, predict_southcoltonbridge, predict_tupperlakenorth, predict_tupperhorseshoe, predict_raquette, predict_pinestreet, predict_piercefield, predict_norwood, predict_norfolk, predict_norfolksm, predict_mountain, predict_mirror, predict_massena, predict_longlake, predict_lehman, predict_lakeclear, predict_lakeflower, predict_horseshoe, predict_heuvelton, predict_govpark, predict_eelweir, predict_dekalb, predict_copper, predict_colton, predict_briggs, predict_blacklake)

# add with lake/river data
clams_predictions_waterbody <- predictions |> left_join(covariate_data, by = "location")

# plot 
plot_RL = ggplot(data = clams_predictions_waterbody,  aes(x = growth_period, y = Length)) +
  geom_line(aes(y = predicted_length, group = name), linewidth = 1) +
  theme_bw() +
  labs(x = "Growth Period", 
       y = "Length", 
       title = "Growth Curves by Lakes and Rivers",
       caption = stringr::str_wrap("Blue line represents the general Von Bertanfly growth model for river vs lake clams. Black lines represent the growth of each clam in the locality.")) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16), 
        plot.caption = element_text(
        size = 14,     
        hjust = 0.5)) +
  facet_wrap(~body) + 
  geom_smooth()
```

```{r}
norfolksm
plot_RL
```

# Bayesian Algorithm

-   Estimate the parameters of the growth curve for river and lake clams.

```{r}
# Rivers

# get and read data
clams <- read_csv("data/clams_clean.csv")
covariate_data <- read_excel("data/covariate_data.xlsx")
clams <- clams |> left_join(covariate_data, by = "location")

# remove negative values from Coopers Falls, Lake Clear, Mirror Lake
clams <- clams |> 
  filter(growth > 0) |>
  mutate(x = case_when(
    body == "L" ~ 1,
    body == "R" ~ 0)) 

# vb function

vb_function <- function(t, max_l, k, b) {
  max_l * (1 - exp(-k * (t - b)))}

# data for rivers
rivers <- clams |> 
  filter(body == "R") |>
  group_by(name)

# MCMC Algorithm


compute_prior_log <- function(max_l, k, b, sigma2) {
  dnorm(max_l, mean = 0, sd = 1000, log = TRUE) +
  dnorm(k, mean = 0, sd = 1000, log = TRUE) +
  dnorm(b, mean = 0, sd = 1000, log = TRUE) +
  log(1 / sigma2)
}

compute_likelihood_log <- function(max_l, k, b, sigma2) {
  dnorm(rivers$Length,
        mean =  vb_function(rivers$growth_period, max_l, k, b),
        sd = sqrt(sigma2), log = TRUE) |>
    sum()
}

compute_posterior_log <- function(max_l, k, b, sigma2) {
  compute_prior_log(max_l, k, b, sigma2) + 
    compute_likelihood_log(max_l, k, b, sigma2)
}


## set-up the chain
niter <- 10000

# set up three empty vectors, one for each parameter
max_l_store <- rep(NA, 10000)
b_store <- rep(NA, 10000)
k_store <- rep(NA, 10000)
sigma2_store <- rep(NA, 10000)
  
  
# define a starting value
max_l_store[1] <- 1.2
k_store[1] <- 0.3183
b_store[1] <- -0.8749
sigma2_store[1] <- 30
  
  
for (i in 2:niter) {
    
     ## start with max_l
  max_l_current <- max_l_store[i - 1]

  proposal_sd <- 3

  max_l_proposal <- rnorm(1, mean = max_l_current, sd = proposal_sd)

  alpha_l <- min(1, exp(compute_posterior_log(max_l_proposal, k_store[i-1], b_store[i-1], sigma2_store[i-1]) - compute_posterior_log(max_l_current, k_store[i-1], b_store[i-1], sigma2_store[i-1])))

  max_l_store[i] <- sample(c(max_l_proposal, max_l_current), size = 1,
                           prob = c(alpha_l, 1 - alpha_l))


  ## start K

  k_current <- k_store[i - 1]

  proposal_sd <- 0.05

  k_proposal <- rnorm(1, mean = k_current, sd = proposal_sd)


  alpha_k <- min(1, exp(compute_posterior_log(max_l_store[i], k_proposal, b_store[i-1], sigma2_store[i-1]) - compute_posterior_log(max_l_store[i], k_current, b_store[i-1], sigma2_store[i-1])))

  k_store[i] <- sample(c(k_proposal, k_current), size = 1,
                       prob = c(alpha_k, 1 - alpha_k))


  # start B

  b_current <- b_store[i - 1]

  proposal_sd <- 0.1

  b_proposal <- rnorm(1, mean = b_current, sd = proposal_sd)

  alpha_b <- min(1, exp(compute_posterior_log(max_l_store[i], k_store[i], b_proposal, sigma2_store[i-1]) - compute_posterior_log(max_l_store[i], k_store[i], b_current, sigma2_store[i-1])))

  b_store[i] <- sample(c(b_proposal, b_current), size = 1,
                       prob = c(alpha_b, 1 - alpha_b))

  # start sigma2

  sigma2_current <- sigma2_store[i-1]
  sigma2_current_log <- sigma2_current |> log()

  proposal_sd <- 0.2

  sigma2_proposal_log <- rnorm(1, mean = sigma2_current_log, sd = proposal_sd)
  sigma2_proposal <- exp(sigma2_proposal_log)

  alpha_sigma2 <- min(1, exp(compute_posterior_log(max_l_store[i], k_store[i], b_store[i], sigma2_proposal) + log(sigma2_proposal) - compute_posterior_log(max_l_store[i], k_store[i], b_store[i], sigma2_current) - log(sigma2_current)))

  sigma2_store[i] <- sample(c(sigma2_proposal, sigma2_current), size = 1,
                            prob = c(alpha_sigma2, 1 - alpha_sigma2))
}
  
  
plot_df_river <- tibble(iter = 1:niter, max_l = max_l_store, k = k_store, b = b_store,
                  sigma2_store) |> filter(iter > 2000)

# max_l plots
river_maxl_line = ggplot(data = plot_df_river, aes(x = iter, y = max_l)) +
  geom_line() +
  labs(x = "iter", y = "max_l",
       title = "max_l by Iteration for River Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))


river_maxl_hist = ggplot(data = plot_df_river, aes(x = max_l)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15)  +
  labs(x = "max_l", y = "frequency",
       title = "Histogram of max_l for River Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

# k plots
river_k_line = ggplot(data = plot_df_river, aes(x = iter, y = k)) +
  geom_line() +
  labs(x = "iter", y = "k",
       title = "K by Iteration for River Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))


river_k_hist = ggplot(data = plot_df_river, aes(x = k)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15) +
  labs(x = "k", y = "frequency",
       title = "Histogram of k for River Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

# b plots
river_b_line = ggplot(data = plot_df_river, aes(x = iter, y = b)) +
  geom_line()

river_b_hist = ggplot(data = plot_df_river, aes(x = b)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15)

# sigma2 plots
river_sigma2_line = ggplot(data = plot_df_river, aes(x = iter, y = sigma2_store)) +
  geom_line()

river_sigma2_hist = ggplot(data = plot_df_river, aes(x = sigma2_store)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15) +
  labs(x = "Sigma-Squared", y = "frequency",
       title = "Histogram of Sigma-Squared for Lake Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))
```

```{r}
# Lakes

# get and read data
clams <- read_csv("data/clams_clean.csv")
covariate_data <- read_excel("data/covariate_data.xlsx")
clams <- clams |> left_join(covariate_data, by = "location")

# remove negative values from Coopers Falls, Lake Clear, Mirror Lake
clams <- clams |> 
  filter(growth > 0) |>
  mutate(x = case_when(
    body == "L" ~ 1,
    body == "R" ~ 0)) 

# vb function

vb_function <- function(t, max_l, k, b) {
  max_l * (1 - exp(-k * (t - b)))}

# data for lakes
lakes <- clams |> filter(body == "L") |> group_by(name)


compute_prior_log <- function(max_l, k, b, sigma2) {
  dnorm(max_l, mean = 0, sd = 1000, log = TRUE) +
  dnorm(k, mean = 0, sd = 1000, log = TRUE) +
  dnorm(b, mean = 0, sd = 1000, log = TRUE) +
  log(1 / sigma2)
}

compute_likelihood_log <- function(max_l, k, b, sigma2) {
  dnorm(lakes$Length,
        mean =  vb_function(lakes$growth_period, max_l, k, b),
        sd = sqrt(sigma2), log = TRUE) |>
    sum()
}

compute_posterior_log <- function(max_l, k, b, sigma2) {
  compute_prior_log(max_l, k, b, sigma2) + 
    compute_likelihood_log(max_l, k, b, sigma2)
}


## set-up the chain
niter <- 10000

# set up three empty vectors, one for each parameter
max_l_store <- rep(NA, 10000)
b_store <- rep(NA, 10000)
k_store <- rep(NA, 10000)
sigma2_store <- rep(NA, 10000)
  
  
# define a starting value
max_l_store[1] <- 1.2
k_store[1] <- 0.3183
b_store[1] <- -0.8749
sigma2_store[1] <- 30
  
  
for (i in 2:niter) {
    
     ## start with max_l
  max_l_current <- max_l_store[i - 1]

  proposal_sd <- 3

  max_l_proposal <- rnorm(1, mean = max_l_current, sd = proposal_sd)

  alpha_l <- min(1, exp(compute_posterior_log(max_l_proposal, k_store[i-1], b_store[i-1], sigma2_store[i-1]) - compute_posterior_log(max_l_current, k_store[i-1], b_store[i-1], sigma2_store[i-1])))

  max_l_store[i] <- sample(c(max_l_proposal, max_l_current), size = 1,
                           prob = c(alpha_l, 1 - alpha_l))


  ## start K

  k_current <- k_store[i - 1]

  proposal_sd <- 0.05

  k_proposal <- rnorm(1, mean = k_current, sd = proposal_sd)


  alpha_k <- min(1, exp(compute_posterior_log(max_l_store[i], k_proposal, b_store[i-1], sigma2_store[i-1]) - compute_posterior_log(max_l_store[i], k_current, b_store[i-1], sigma2_store[i-1])))

  k_store[i] <- sample(c(k_proposal, k_current), size = 1,
                       prob = c(alpha_k, 1 - alpha_k))


  # start B

  b_current <- b_store[i - 1]

  proposal_sd <- 0.1

  b_proposal <- rnorm(1, mean = b_current, sd = proposal_sd)

  alpha_b <- min(1, exp(compute_posterior_log(max_l_store[i], k_store[i], b_proposal, sigma2_store[i-1]) - compute_posterior_log(max_l_store[i], k_store[i], b_current, sigma2_store[i-1])))

  b_store[i] <- sample(c(b_proposal, b_current), size = 1,
                       prob = c(alpha_b, 1 - alpha_b))

  # start sigma2

  sigma2_current <- sigma2_store[i-1]
  sigma2_current_log <- sigma2_current |> log()

  proposal_sd <- 0.2

  sigma2_proposal_log <- rnorm(1, mean = sigma2_current_log, sd = proposal_sd)
  sigma2_proposal <- exp(sigma2_proposal_log)

  alpha_sigma2 <- min(1, exp(compute_posterior_log(max_l_store[i], k_store[i], b_store[i], sigma2_proposal) + log(sigma2_proposal) - compute_posterior_log(max_l_store[i], k_store[i], b_store[i], sigma2_current) - log(sigma2_current)))

  sigma2_store[i] <- sample(c(sigma2_proposal, sigma2_current), size = 1,
                            prob = c(alpha_sigma2, 1 - alpha_sigma2))
}
  
  
plot_df_lakes <- tibble(iter = 1:niter, max_l = max_l_store, k = k_store, b = b_store,
                  sigma2_store) |> filter(iter > 2000)

# max_l plots
lakes_maxl_line = ggplot(data = plot_df_lakes, aes(x = iter, y = max_l)) +
  geom_line() +
  labs(x = "iter", y = "max_l",
       title = "max_l by Iteration for Lake Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

lakes_maxl_hist = ggplot(data = plot_df_lakes, aes(x = max_l)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15) +
  labs(x = "max_l", y = "frequency",
       title = "Histogram of max_l for Lake Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16)) 

# k plots
lakes_k_line = ggplot(data = plot_df_lakes, aes(x = iter, y = k)) + 
  geom_line() +
  labs(x = "iter", y = "k",
       title = "K by Iteration for Lake Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

lakes_k_hist = ggplot(data = plot_df_lakes, aes(x = k)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15)+
  labs(x = "k", y = "frequency",
       title = "Histogram of k for Lake Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

# b plots
lakes_b_line = ggplot(data = plot_df_lakes, aes(x = iter, y = b)) +
  geom_line()

lakes_b_hist = ggplot(data = plot_df_lakes, aes(x = b)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15)

# sigma2 plots
lakes_sigma2_line = ggplot(data = plot_df_lakes, aes(x = iter, y = sigma2_store)) +
  geom_line()

lakes_sigma2_hist = ggplot(data = plot_df_lakes, aes(x = sigma2_store)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15) +
  labs(x = "Sigma-Squared", y = "frequency",
       title = "Histogram of Sigma-Squared for Lake Clams") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))
```

```{r}
results_river <- plot_df_river |>
  summarise(median_l = median(max_l),
            median_k = median(k),
            median_b = median(b),
            median_sigma2 = median(sigma2_store))|>
  mutate(body = "River")

results_lake <- plot_df_lakes |> 
  summarise(median_l = median(max_l),
            median_k = median(k),
            median_b = median(b),
            median_sigma2 = median(sigma2_store)) |>
              mutate(body = "Lake")
  

results <- bind_rows(results_lake, results_river)

pander(results, caption = "Parameter Estimates for Lakes and Rivers")
```

```{r}
# graphs 
plot_df_lakes <- plot_df_lakes |> mutate(body = "Lake")
plot_df_river <- plot_df_river |> mutate(body = "River")

graphs_df <- bind_rows(plot_df_lakes, plot_df_river) 

maxl_line = ggplot(data = graphs_df, aes(x = iter, y = max_l)) + 
  geom_line() +
  facet_wrap(~body) +
  labs(x = "iter", y = "max_l",
       title = "Lake vs. River: Value of max_l through Each Iteration") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

maxl_hist = ggplot(data = graphs_df, aes(x = max_l, fill = body)) +
  scale_fill_manual(values = c("Lake" = "steelblue", "River" = "navy")) + 
  geom_histogram(bins = 15, colour = "black") +
  labs(x = "max_l", y = "frequency",
       title = "Lake vs. River: Histogram of max_l Values") +   
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))


k_line = ggplot(data = graphs_df, aes(x = iter, y = k)) + 
  geom_line() +
  facet_wrap(~body) +
  labs(x = "iter", y = "k",
       title = "Lake vs. River: Value of k through Each Iteration") + 
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))

k_hist = ggplot(data = graphs_df, aes(x = k, fill = body)) +
  scale_fill_manual(values = c("Lake" = "steelblue", "River" = "navy")) + 
  geom_histogram(bins = 15, colour = "black") +
  labs(x = "k", y = "frequency",
       title = "Lake vs. River: Histogram of k Values") +   
  theme_minimal(base_size = 24, base_rect_size = 50) +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
        axis.text.x  = element_text(size = 12),
        plot.subtitle = element_text(size = 16),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16))
```

```{r, out.width= "70%", fig.align='center'}
maxl_hist
```
River clams have larger maximum lengths than lake clams. This means that river clams grow longer. 

```{r, out.width= "70%", fig.align='center'}
k_hist
```
Lake clams have a larger k. This means that lake clams reach their maximum length quicker than river clams. 

```{r, out.width= "70%", fig.align='center'}
lakes_maxl_line
```
The values for max_l fluctuate around a similar value, meaning that the algorithm converges. 

# Take Aways

- River clams grow to longer overall lengths than lake clams. 

- Lake clams reach their asymptotic length twice as fast as river clams. 

- River water is less acidic than lake water. Clam shells are made out of calcium carbonate, which dissolve in acidic water. Differences in acidity between lake and rivers may partially explain the differences in shell length and growth rate. 

- River current promotes long-term growth by delivering organic particles for filter feeding and maintaining higher dissolved oxygen levels, explaining the longer length.

# Future Direction

- Improve the algorithm to account for uncertainty in the timing of the initial observed growth period relative to actual age.

- Expand the algorithm to estimate parameters for every location. 

# References and Acknowledgements

- Johnson, A. A., Ott, M. Q., & Dogucu, M. (2022). Bayes rules! An introduction to applied Bayesian modeling. https://www.bayesrulesbook.com/

- Windsland, K., Hvingel, C., Nilssen, E. M., & Sundet, J. H. (2013). Evaluation of von Bertalanffy growth curves for the introduced red king crab (Paralithodes camtschaticus) in Norwegian waters. Fisheries Research, 145, 15–21. https://doi.org/10.1016/j.fishres.2013.03.003

- Thank you to Dr. Judith Nagel-Myers for letting me use her data. 