---
title: "Write Up"
author: "Emma Deering"
format: html
---

**Introduction to Data:**

- Clams grow through a process called accretion, where every growing period their shells expand by adding new material to the outer edge. This creates a new visible ring around the shell. The number and thickness of these rings depend on a variety of factors, such as climate, food availability, and water conditions. 

- These rings are preserved in fossils, allowing geologists and paleontologists to make conclusions regarding fossilized clams ages and environmental conditions. Specifically, whether ADK lakes or rivers impact growth of clams. 

- This data looks at clam growth per growth period for fossilized clams in bodies of water across the Adirondacks and Northern New York. 

- The data set contains the following useful variables:
  Location: The lake or river that the clam lived in
  Growth Period: Indicates stage of growth. 1 growth period represents 1 year of life.
  Length: Overall length of shell
  Growth: Total growth during that growth period
  Body: Indicates if water body was a lake or river. 
  
- The data used for this project was recorded by Dr. Judith Nagel-Myers of the Geology Department. She collected 151 valves from 17 different localities. 
  
**Introduction to Growth Model:**

- For this study, the Von Bertanfly Growth model was used. The Von Bertanfly Growth model is a mathematical model that is widely accepted and used to interpret and understand organisms development over time.

- Given the growth type of clams, this model will best illustrate the data since it is designed to model organism development, especially growth that decreases with time, such as seen in clams. It models length as a function of age.

- The model looks like this: 

L(t) = L x (1 - exp(-k(t-t0)))

where 
- L(t) is mean length at time t
- L is the asymptotic length
- k is the rate
- t is time
- t0 is time where length is 0

- The goal of using the growth model is to understand growth patterns in the fossilized clams in order to draw conclusions about how clams grow in different aquatic environments.

*Add growth image maybe from one of the clams and describe meaning*
```{r}
norfolksm <- ggplot(predict_norfolksm,
  aes(x = growth_period, y = Length)) +
  geom_point() +
  geom_line(aes(y = predicted_length),
            color = "red", linewidth = 2) +
  theme_bw() +
  labs(x = "Age", y = "Growth",
       title = "Von Bertanfly Growth Model")

norfolksm
```

- This image illustrates an example of the Von Bertanfly Growth Model. Here, the model accurately represents the growth of organisms the grow faster in youth and decrease with age. For every year, the rate of growth decreases, giving the model a logarithmic look. 

**Introduction to Bayesian Statistics:**

- Many of the clams may be broken, so determining the earliest date of growth may cause issue. To work around this, I will use Bayesian statistics since it is more flexible than other traditional statistical methods

- Bayesian statistics works by using our prior data with our new information to construct our posterior knowledge. As we get more and more information, we can continue to add to what we know and adjust the model. 

- This method of thinking better suits the data where clams may not all be starting at the same growth period. Instead, as we get new information with each growth period, we will be able adjust the model slightly till it is completely refined.

- Maybe insert that image from the textbook explaining. 

**Data Exploration:**

Here we explore the data for initial relationships, understand the data and summary of the data. 

```{r}
ggplot(data = clams_predictions_waterbody, aes(x = growth_period, y = lag_length)) +
  geom_point() +
  geom_smooth() +
  theme_bw()
```
Here, we see that as a clam ages, the length increases. However, it increases rapidly in early life, and the grate rate slowly decreases later in life. This shows that the Van Bertanfly Growth Model is the proper model for this data. 

```{r}
max_length <- clams_predictions_waterbody |> group_by(name) |> mutate(max = max(Length))

ggplot(data = max_length, aes(x = max)) + 
  geom_histogram(colour = "skyblue4", fill = "skyblue1", bins = 15) +
  theme_bw() +
  labs(x = "Overall Length", 
       y = "Frequency", 
       title = "Histogram of Total Length")
```
For this study, we used a variety of overall sizes of clams. Since length increases with age, as shown above, we can see we use a normal distribution of different aged clams, where the majority are middle ages with a few young and old samples.

```{r}
ggplot(data = max_length, aes(y = max, x = body)) + 
  geom_boxplot(colour = "skyblue4", fill = "skyblue1") +
  theme_bw() +
  labs(x = "Type of Water Body", 
       y = "Total Length", 
       title = "Boxplot of Total Length by Lakes & Rivers")
```
It appears length of the clams are related to the body of water that it originates in, specifically lakes or rivers. It appears rivers have  longer clams than clams from lakes. This likely has something to do with river conditions, including food content, rivers, and acidity.

```{r}
ggplot(data = clams_predictions_waterbody,  aes(x = growth_period, y = Length)) +
  geom_line(aes(y = predicted_length, group = name), linewidth = 1) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length", title = "Growth Curves of All Clams") +
  geom_smooth()
```
Here, we used the Von Bertanfly Growth model on all the clams in the data set. Here we see that the model well illustrates the growth rates of clams. Clams tend to have faster growth rates at the beginning of their life, and slows as they get older. 


```{r}
ggplot(data = clams_predictions_waterbody,  aes(x = growth_period, y = Length)) +
  geom_line(aes(y = predicted_length, group = name), linewidth = 1) +
  theme_bw() +
  labs(x = "Growth Period", y = "Length", title = "Growth Curves by Lakes and Rivers") +
  facet_wrap(~body) + 
  geom_smooth()
```
Here, we have divided the growth curves by bodies of water, specifically lakes and rivers. Rivers have much faster and longer growth rates than lakes.  

Overall, as we move into our analysis, we can see that growth rate is affected by age making this data a good candidate for the Von Bertanfly Growth model, but more interestingly the body of water that the clams orginiates from appears to be an important factor in its growth.
